---
title: "index"
author: "Julia Stuart"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Add libraries needed
```{r library, warning=FALSE, message=FALSE}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)

```

Read in data from: Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r import data}

esc <- read.csv(file = "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = FALSE)

```
#Data Tidying

##Calculate annual, regional total escarpment

* `separate` the date into Year-Month-Date

* `group_by` year, species, and region and `summarise`

* `filter` for salmon species


```{r cleaning data, echo=FALSE}

annual_esc <- esc %>%
  separate(sampleDate, into = c("year","month","day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>%  #change year from <chr> to <dbl>
  group_by(SASAP.Region,Species,year) %>% 
  summarise(escarpment = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook","Sockeye","Pink","Chum","Coho"))
#head(annual_esc)
```

```{r interactive table, echo=FALSE}

datatable(annual_esc)
```

#Static Plots

```{r ggplot}
names(annual_esc)

ggplot(data = annual_esc, mapping = aes(x = Species, y = escarpment)) +
  geom_col()

#everything in the aes call is related to a column in your dataset, therefore
#you need to 

ggplot(annual_esc, aes(x = Species, y = escarpment, fill=SASAP.Region)) +
  geom_col()





```

Now say we're interested in just Kodiak species
timeseries of Kodiak by species

```{r ggplot plus}

ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), aes(year,escarpment, color = Species)) + 
  geom_line()

#now make it look nicer

my_theme <- theme_bw()+
  theme(plot.title = element_text(hjust=0.5))

ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = escarpment, color = Species)) + 
    geom_line() +
    geom_point(shape = 12, size=.5) +
    ylab("Escapement") +
    ggtitle("Kodiak Salmon Escapement") +
    ggtitle("Kodiak Salmon") +
  my_theme
```
Plot timeseries of escapement by species for all regions

```{r}

```


#Interactive Maps





