---
title: "index"
author: "Julia Stuart"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
If you don't want to show ANY code, above set echo = FALSE 
Add libraries needed
```{r library, warning=FALSE, message=FALSE}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(ncdf4)
library(rnoaa)

```

Read in data from: Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r import data}

esc <- read.csv(file = "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = FALSE)

```
#Data Tidying

##Calculate annual, regional total escarpment

* `separate` the date into Year-Month-Date

* `group_by` year, species, and region and `summarise`

* `filter` for salmon species


```{r cleaning data, echo=FALSE}

annual_esc <- esc %>%
  separate(sampleDate, into = c("year","month","day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>%  #change year from <chr> to <dbl>
  group_by(SASAP.Region,Species,year) %>% 
  summarise(escarpment = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook","Sockeye","Pink","Chum","Coho"))
#head(annual_esc)
```

```{r interactive table, echo=FALSE}

datatable(annual_esc)
```

#Static Plots

```{r ggplot}
names(annual_esc)

ggplot(data = annual_esc, mapping = aes(x = Species, y = escarpment)) +
  geom_col()

#everything in the aes call is related to a column in your dataset, therefore
#you need to 

ggplot(annual_esc, aes(x = Species, y = escarpment, fill=SASAP.Region)) +
  geom_col()





```

Now say we're interested in just Kodiak species
timeseries of Kodiak by species

```{r ggplot plus}

ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), aes(year,escarpment, color = Species)) + 
  geom_line()

#now make it look nicer

my_theme <- theme_bw()+
  theme(plot.title = element_text(hjust=0.5))

ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = escarpment, color = Species)) + 
    geom_line() +
    geom_point(shape = 12, size=.5) +
    ylab("Escapement") +
    ggtitle("Kodiak Salmon Escapement") +
    ggtitle("Kodiak Salmon") +
  my_theme
```
Plot timeseries of escapement by species for all regions

```{r facet wrapping, fig.height=10, fig.width=6}

ggplot(annual_esc, aes(year, escarpment, color=Species)) +
  geom_line()+
  geom_point(shape = 12, size=.5)+
  facet_wrap(~SASAP.Region, scales= "free", ncol=2)+
  theme_bw()

ggsave("region_e.png")

```


#Interactive Maps
Find unique locations with lat-long values
```{r unique locations}

locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()

datatable(locations)

```

Creating leaflet map!

```{r leaflet}
#leaflet uses the pipe, not the plus
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~Latitude, popup = ~Location)
```



#R OpenSci Demo

Read in buoy data
```{r read in buoy data}
bd <- buoy(dataset= "cwind", buoyid = 46080, year = 2018, datatype = "c")
```
Make it a real dataframe based on the buoy object
```{r make a real dataframe}
wind_speed <- data.frame(time = as.POSIXct(bd$data$time), speed = bd$data$wind_spd)
#wrap "time" values in as.POSIXct to handle time elements better
#different than string, numeric, or factor

```

```{r plot windspeed}
ggplot(wind_speed, aes(time, speed))+
  geom_point(shape=5, size=1)

```
^This was for the physical scientists
Now, for the Biologists!

```{r mapr and spocc}
#install.packages("mapr", dependencies = TRUE)
library(mapr)
library(spocc)

spp <- c("Ursus arctos", "Ursus maritimus","Ursus americanus kermodei")
bears <- occ(query = spp, from = 'gbif', has_coords = TRUE, limit = 50)
map_leaflet(bears, color = c("brown", "blue","pink"), size = 1)
```

CAN YOU DO THIS FOR MOSS??

```{r YOU CAN DO MOSS}
spp2 <- c("Hylocomium splendens", "Tomentypnum nitens","Sphagnum fuscum")
moss <- occ(query = spp2, from = 'gbif', has_coords = TRUE, limit = 500)
map_leaflet(moss, color = c("#498A8B", "#7E8D60","#838E8E"), size = 1, alpha =1)
```
There is a package called striptime that can do cool stuff with dates.
I'm writing this with the hope that I'll never have to use it. 

#Net CDF files
##Read in data from a file
```{r import parts of file needed}
setwd("~/Desktop/ADC_training/ADC_test")
nc <- nc_open("~/Desktop/ADC_training/WG2013CTD.nc")
print(nc) #use this for an overview of what is in the profile
```
NetCDF files are very flexible, and that is both a good and a bad thing
```{r assign variables to vector, see dimensions}
#apparently this will help us know how the matrix is oriented
#this is also listed in the metadata
vars <- attributes(nc$var)$names
vars

dims <- attributes(nc$dim)$names
#this makes it easier to look at the names of the data you might want
```
If we were doing this "for real", there's some other stuff we could pull 
Note: these values in the box above didn't "pop up" when I used the $; I typed what was in the tutorial

```{r make salinity & time values}
salinity <- ncvar_get(nc, "sal")
time <- ncvar_get(nc, "time")
depth <- ncvar_get(nc, "z")
```

##Reformat the output
Coerce this into a dataframe
```{r fix the time}

time <- as.POSIXct((time + 719529)*86400, origin = "1970-01-01", tz = "UTC")
#this would be a good function for the mytools package 
```

```{r make salinity dataframe}
salinity_data <- as.data.frame(salinity)
#Now assign column names to depth values
names(salinity_data) <- as.character(depth)
#Reshape the data and add time
salinity_data_long <- salinity_data %>% 
  mutate(time = time) %>% 
  gather(key = "depth", value = "salinity", -time) %>% 
  mutate(depth = as.numeric(depth))

#probably should retain a profile ID, should've grabbed from netCDF

```

##Plots

```{r}
ggplot(salinity_data_long, aes(x = time, y = depth, fill = salinity)) +
    geom_raster() +
    theme_bw() +
    ylab("Depth (m)") +
    xlab("") +
    scale_fill_continuous(low = "gray", high = "red", name = "Salinity (psu)")

ggplot(salinity_data_long, aes(x = salinity,
                          y = depth,
                          group = time,
                          color = time)) +
    geom_line(size = .1) +
    scale_y_reverse() +
    theme_bw() +
    ylab("Depth (m)") +
    xlab("Salinity (psu)") +
    theme(legend.title = element_blank())
```
If you are working with rester data, you can also use the `raster` package


