---
title: "tidy_adc"
author: "Julia Stuart"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Keeping the top setup chunk^ so you could change echo defaults for the whole doc

Life advice from ADC: knowing when to commit is a very personal thing

Include all library calls at then beginning of the script
```{r library, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
```
If you need to use a masked function, package_name::function()


#Data Cleaning

Read in datafile:

```{r import data}
catch <- read.csv(file = "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

#called byerlySalmonByRegion.csv

head(catch)
```

Possible alt code if that doesn't work:

Note, for this Markup I added ,eval=FALSE to the curly r brackets

That means that the code will show up, but not run

```{r import data alt method,eval=FALSE}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```

##Pipe Operators

%>% the keyboard shortcut is command+shift+M 

Added warning=FALSE for the warning about variables? 

```{r data piping to clean, warning=FALSE}
names(catch)

catch_long <- catch %>%
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>%
  gather(key = "species", value = "catch", -Year, -Region)
#keys are column names (headings), values are values contained in the columns
#the opposite of gather is spread

#OR 

#catch_clean <- catch %>%
  #select(-All, -notesRegCode)

head(catch_long)
  

```

In the above code chunk, we 
*Removed two colums
*Converted from wide data using gather for species counts



#Split-Apply-Combine

#join jawns

