---
title: "tidy_adc"
author: "Julia Stuart"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Keeping the top setup chunk^ so you could change echo defaults for the whole doc

Life advice from ADC: knowing when to commit is a very personal thing

Include all library calls at then beginning of the script
```{r library, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
```
If you need to use a masked function, package_name::function()


#Data Cleaning

Read in datafile:

```{r import data}
catch <- read.csv(file = "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

#called byerlySalmonByRegion.csv

head(catch)
```

Possible alt code if that doesn't work:

Note, for this Markup I added ,eval=FALSE to the curly r brackets

That means that the code will show up, but not run

```{r import data alt method,eval=FALSE}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```

##Pipe Operators

%>% the keyboard shortcut is command+shift+M 

Added warning=FALSE for the warning about variables? 

This is the warning: attributes are not identical across measure variables;
they will be dropped

```{r data piping to clean, warning=FALSE}
names(catch)

catch_long <- catch %>%
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>%
  gather(key = "species", value = "catch", -Year, -Region)
#keys are column names (headings), values are values contained in the columns
#the opposite of gather is spread

#OR 

#catch_clean <- catch %>%
  #select(-All, -notesRegCode)

head(catch_long)
  

```

In the above code chunk, we 
*Removed two colums
*Converted from wide data using gather for species counts



```{r add columns, remove problems}

#issue: counts are in thousands of fish. They did not catch 7 fish in all of Alaska

catch_clean <- catch_long %>%
  rename(catch_thousands = catch) %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I",1,catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands))%>%
  mutate(catch = catch_thousands*1000)

names(catch_clean)
summary(catch_clean)
#general tip: the thing you're creating is on the left, the old one is on the right

```
Uh oh! I got a warning message-- non numeric operators -- in the last pipe.

The code below will locate the problem

And then I return to the add columns, remove problems section to fix.

```{r testing chunk, eval=FALSE}

#first I "commented" the section of code that didn't work above

test_catch <- as.integer(catch_clean$catch_thousands)
#NAs introduced by coercion
#but you can locate those NAs:
i <- which(is.na(test_catch) == T)
#now print the location of the i 
catch_clean[i,]
#telling us that in row 401, catch_thousands is a letter not a number

#now that that's solved, we will return to the pipe chunk
#and add another line to the pipe operator
#see:ifelse statement
#after the ifelse statement has run, re-run the test_catch and i commands
#then, in the global enviro, see what the value of i is
#should change to interger(empty)
#after that, add another line to the pipe O to mutate catch_thous to integers
#use head to check if it IS indeed turned into an integer 

```

How to fix your character flaws: see additions to add columns, remove problems






#Split-Apply-Combine

#join jawns

